Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over convex domains and domains with shellable triangulations

### 2024-10-04 draft.tex
further polishment

>>>>

\section{{Nonlinear and linear potentials, the Poincar\'e--Friedrichs and Poincar\'e inequalities, and {analytical} Poincar\'e--Friedrichs constants in the gradient case}} \label{section:poincare}

\mwl{
In this section, we wish to explain why we employ linear potential operators to estimate the Poincar\'e--Friedrichs constants and we also address the difference of the Poincar\'e--Friedrichs inequality~\eqref{math:PF_gradient_2} with the Poincar\'e inequality where the mean value of $u$ is fixed to zero. We then review Poincar\'e--Friedrichs inequalities in the gradient case setting of Section~\ref{section:intro_grad}, with the emphasis on know results on analytical upper bounds for the constants. In particular, we take the opportunity to discuss bounded convex domains as a special case. 
This survey serves as a building block in constructing computable constants over triangulated domains in a combinatorial way below but we believe it is also of independent interest.

It is generally easier to obtain upper bounds Constructing and analyzing a potential operator is generally easier if the potential is linear. 
Constructing linear potential operators is generally easier 
We will extend that construction of the potential when we address Poincar\'e--Friedrichs inequalities for the curl and divergence operators.

% Our particular notion of Poincar\'e--Friedrichs inequality, which we interpret as an upper bound for the operator norm of a generalized inverse of the gradient is more natural for our discussion because later sections will develop analogous inequalities for the curl and the divergence operators. 
}

In this section, we wish to explain why we employ linear potential operators to estimate the Poincar\'e--Friedrichs constants and we also address the difference of the Poincar\'e--Friedrichs inequality~\eqref{math:PF_gradient_2} with the Poincar\'e inequality where the mean value of $u$ is fixed to zero. We then review Poincar\'e--Friedrichs inequalities in the gradient case setting of Section~\ref{section:intro_grad}, with the emphasis on know results on analytical upper bounds for the constants. In particular, we take the opportunity to discuss convex domains as a special case. This survey serves as a building block in constructing computable constants over triangulated domains in a combinatorial way below but we believe it is also of independent interest.

\subsection{Nonlinear and linear potentials}

In the Banach setting $p \neq 2$, finding a norm-minimizing potential such as $\Phi_{\grad}( \bff )$ of~\eqref{math:pot_grad_def} for a given gradient $\bff$ is a nonlinear operation. Its existence may actually not be guaranteed unless $1 < p < \infty$. 
However, any \textit{linear} potential operator $\overline\Phi_{\grad} : \nabla W^{1,p}(\Omega) \rightarrow L^{p}(\Omega)$
satisfying $\nabla \overline\Phi_{\grad}( \bff ) = \bff$ for any $\bff \in \nabla W^{1,p}(\Omega)$ has the operator norm that \cye{gives} the upper bound 
\begin{align}\label{math:PF_P}
    \mwl{C_{\grad,\Omega,p} = } 
    \max\limits_{ u \in W^{1,p}(\Omega) \setminus \bbR } 
    \frac{ \| \Phi_{\grad}(\nabla u) \|_{L^{p}(\Omega)} }{ \| \nabla u \|_{\fooL^{p}(\Omega)} }
    \leq 
    \max\limits_{ u \in W^{1,p}(\Omega) \setminus \bbR } 
    \frac{ \| \overline\Phi_{\grad}(\nabla u) \|_{L^{p}(\Omega)} }{ \| \nabla u \|_{\fooL^{p}(\Omega)} } \cye{=: C_{\varnothing,\Omega,p}}
    .
\end{align}
This explains our interest in the construction of {\em linear potential operators}.

\subsection{Poincar\'e--Friedrichs and Poincar\'e inequalities}

\cye{In the setting of Section~\ref{section:intro_grad}, when $\Omega$ is of finite measure,} a natural \cye{(and unique, since the gradient potential constraint $\nabla \overline \Phi_{\grad}( \bff ) = \bff$ determines $\overline \Phi_{\grad}( \bff )$ up to a constant)} choice \cye{for a linear potential operator} is the average-free potential operator $\overline\Phi_{\grad} : \nabla W^{1,p}(\Omega) \rightarrow L^{p}(\Omega)$ that is prescribed by
\begin{align*}
    \overline\Phi_{\grad}( \nabla u ) \cye{:}= u - u_{\Omega},
    \quad 
    \forall 
    u \in W^{1,p}(\Omega)
    ,
\end{align*}
where $u_\Omega$ denotes the average of $u$ over the domain $\Omega$, that is,
\begin{align*}
    u_{\Omega} := \vol(\Omega)^{-1} \int_{\Omega} u(x) \;dx.
\end{align*}
This potential operator can also implicitly be defined by
\begin{align*}
    \overline \Phi_{\grad}( \bff ) 
    := 
    \argmin\limits_{ \substack{ u \in W^{1,p}(\Omega) \\ \int_\Omega u = 0 } } \| \nabla u - \bff \|_{\fooL^{p}(\Omega)},
    \quad 
    \forall 
    \bff \in \nabla W^{1,p}(\Omega)
    .
\end{align*} 
Its operator norm is the best constant $C_{\varnothing,\Omega,p} > 0$ in the {\em Poincar\'e inequality} (or also Poincar\'e--Wirtinger inequality)
\begin{align}\label{math:P_gradient}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    C_{\varnothing,\Omega,p} \| \nabla u \|_{\fooL^{p}(\Omega)},
    \quad 
    \forall 
    u \in W^{1,p}(\Omega)
    . 
\end{align}
Thus, rephrasing~\eqref{math:PF_P} in terms of~\eqref{math:PF_gradient_2} and~\eqref{math:P_gradient}, 
\begin{align*}
    C_{\grad,\Omega,p} \leq C_{\varnothing,\Omega,p}.
\end{align*}

Towards a converse inequality, 
let us first observe that the average of any $u \in W^{1,p}(\Omega)$ with $p < \infty$ satisfies the bound 
\begin{align} \label{math:aver_norm}
    \| u_\Omega \|_{L^{p}(\Omega)}^{p}
    = 
    \int_{\Omega} \left( \vol(\Omega)^{-1} \int_{\Omega} |u(x)| \;dx \right)^{p}
    \leq 
    \int_{\Omega} \vol(\Omega)^{-1} \int_{\Omega} |u(x)|^{p} \;dx
    = 
    \| u \|_{L^{p}(\Omega)}^{p}
    .
\end{align}
Here, we have used H\"older's or Jensen's inequality. 
In the case $p = \infty$, any $u \in L^{\infty}(\Omega)$ satisfies $\| u_\Omega \|_{L^{\infty}(\Omega)} \leq \| u \|_{L^{\infty}(\Omega)}$. 
We conclude that taking the average is a projection within Lebesgue spaces with unit norm. 
The triangle inequality now shows that 
\begin{align*}
    \| u - u_\Omega \|_{L^{p}(\Omega)} 
    \leq
    2
    \| u \|_{L^{p}(\Omega)},
    \quad 
    \forall
    u \in L^{p}(\Omega)
    .
\end{align*}
\begin{align*}
    \color{Emerald}
    \| u - u_\Omega \|_{L^{p}(\Omega)}^{p}
    \leq 
    2^{p-1} 
    \| u \|_{L^{p}(\Omega)}^{p}
    +
    2^{p-1} 
    \| u_\Omega \|_{L^{p}(\Omega)}^{p}
    \leq 
    2^{p} 
    \| u \|_{L^{p}(\Omega)}^{p}
    .
\end{align*}
\mwl{Via an application of the triangle inequality, }Thus, the Poincar\'e--Friedrichs inequality implies the Poincar\'e inequality with 
\begin{align*}
    C_{\grad,\Omega,p} \leq 2 C_{\varnothing,\Omega,p}.
\end{align*}
In the special case $p=2$, taking the average is an orthogonal projection, 
and so this improves to $\| u - u_\Omega \|_{L^{2}(\Omega)} \leq \| u \|_{L^{2}(\Omega)}$ for any $u \in L^{2}(\Omega)$. 
Hence,
\begin{align*}
    C_{\varnothing,\Omega,2} = C_{\grad,\Omega,2}. 
\end{align*}
\mwl{
This improvement follows from the orthogonality of constant functions to average-free functions in $L^2(\Omega)$,
and it also follows from the projection estimate (see, e.g.,~\cite{xu2003some}).
Stern's generalized projection estimate~\cite[Theorem~4.1,Remark~5.1]{stern2015banach} implies improved estimate for all $L^{p}$ spaces with $1 \leq p \leq \infty$:
since taking the average is a the projection onto the constants functions with unit norm, it now follows that 
}
From~\eqref{math:aver_norm}, Stern's generalized projection estimate~\cite[Theorem~4.1, Remark~5.1]{stern2015banach} implies improved estimate for all $L^{p}$ spaces with $1 \leq p \leq \infty$ as 
\begin{align*}
    \| u - u_\Omega \|_{L^{p}(\Omega)}
    \leq 
    \min\left( 2, 2^{|2/p-1|} \right)
    \| u \|_{L^{p}(\Omega)}
    = 
    2^{|2/p-1|} 
    \| u \|_{L^{p}(\Omega)}
    ,
    \quad 
    \forall 
    u \in L^{p}(\Omega)
    .
\end{align*}
Here, we have used $1 \leq 2^{|2/p-1|} \leq 2$ for $1 \leq p \leq \infty$.
The upper bound is only attained in the limit cases $p = 1$ and $p = \infty$, and the lower bound is only attained if $p = 2$, where we reproduce the improved estimate for the Hilbert space case.
We thus conclude 
\begin{align}\label{math:p_vs_pf}
    C_{\grad,\Omega,p} \leq 2^{|2/p-1|} C_{\varnothing,\Omega,p}.
\end{align}
In summary, our notion of Poincar\'e--Friedrichs constant is equivalent to the notion of Poincar\'e constant in the literature, up to a numerical factor that only depends on $1 \leq p \leq \infty$ and that is at most $2$.

%%%%
